Model

### 2020-02-02 stats/model.Rmd
update model processing code

# Ignore below for now

```{r}
log_like <- bind_rows(
  readRDS("../saves/log_like_chain1.rds")%>%mutate(chain = 1),
  readRDS("../saves/log_like_chain2.rds")%>%mutate(chain = 2),
  readRDS("../saves/log_like_chain3.rds")%>%mutate(chain = 3),
  readRDS("../saves/log_like_chain4.rds")%>%mutate(chain = 4),
  readRDS("../saves/log_like_chain5.rds")%>%mutate(chain = 5)
)

@@

@@

### Pragmatic free model

@@

plot_model_pred <- model_pred%>%
  filter(model == "pragmatic")%>%
  group_by(run, model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = plot_model_pred, aes(x=age+2,y = mean, col = run), size = 1)+
  geom_ribbon(data = plot_model_pred, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = run), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol()+
  scale_fill_ptol()

length(unique(model_pred$iteration))

x <- sample(1:10000, 500)

plot_model_prag__line <- model_pred%>%
  filter(model == "pragmatic", iteration %in% x)

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(data = plot_model_prag__line, aes(x=age+2,y = pred, col = alignment, group = interaction(chain,iteration)),size = .2, alpha = .05)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "glm", method.args = list(family = "binomial"), se = T, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)
```

@@

### Pragmatic noise model

@@

model_pred_prag_noise<- bind_rows(
  readRDS("../saves/model_pred_pragmatic_noise_chain1.rds")%>%mutate(chain = 1),
  readRDS("../saves/model_pred_pragmatic_noise_chain2.rds")%>%mutate(chain = 2),
  readRDS("../saves/model_pred_pragmatic_noise_chain3.rds")%>%mutate(chain = 3),
  readRDS("../saves/model_pred_pragmatic_noise_chain4.rds")%>%mutate(chain = 4),
  readRDS("../saves/model_pred_pragmatic_noise_chain5.rds")%>%mutate(chain = 5)
)

plot_model_pred_prag_noise <- model_pred_prag_noise%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

saveRDS(plot_model_pred_prag_noise, "../saves/plot_model_pred_prag_noise.rds")
plot_model_pred_prag_noise <- readRDS("../saves/plot_model_pred_prag_noise.rds")

plot_model_pred_prag_noise_chain <- model_pred_prag_noise%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item,chain, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

saveRDS(plot_model_pred_prag_noise_chain, "../saves/plot_model_pred_prag_noise_chain.rds")
plot_model_pred_prag_noise_chain <- readRDS("../saves/plot_model_pred_prag_noise_chain.rds")

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_ribbon(data = plot_model_pred_prag_noise, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .3) +
  geom_line(data = plot_model_pred_prag_noise, aes(x=age+2,y = mean, col = model), size = 1, alpha = .5)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)

x <- sample(1:48000, 500)

plot_model_prag_noise_line <- model_pred_prag_noise%>%
  filter(iteration %in% x)

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .4, span = 2)+
  geom_line(data = plot_model_prag_noise_line, aes(x=age+2,y = pred, col = factor(chain), group = interaction(chain,iteration)),size = .2, alpha = .075)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)
```

```{r}
ggsave("../graphs/model_pred_noise_ribbon.pdf", width = 10, height = 4, scale = 1.2)
```

### Global pragmatic free model

model_pred_global_free<- bind_rows(
  readRDS("../saves/model_pred_globalPrag_free_chain1.rds")%>%mutate(chain = 1),
  readRDS("../saves/model_pred_globalPrag_free_chain2.rds")%>%mutate(chain = 2)
)

plot_model_pred_global_free <- model_pred_global_free%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

saveRDS(plot_model_pred_global_free, "../saves/plot_model_pred_global_free.rds")
plot_model_pred_global_free <- readRDS("../saves/plot_model_pred_global_free.rds")

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = plot_model_pred_global_free, aes(x=age+2,y = mean, col = model), size = 1)+
  geom_ribbon(data = plot_model_pred_global_free, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(model + alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```

### Global pragmatic noise model

model_pred_global_noise<- bind_rows(
  readRDS("../saves/model_pred_globalPrag_noise_chain1.rds")%>%mutate(chain = 1),
  readRDS("../saves/model_pred_globalPrag_noise_chain2.rds")%>%mutate(chain = 2)
)

plot_model_pred_global_noise <- model_pred_global_noise%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

saveRDS(plot_model_pred_global_noise, "../saves/plot_model_pred_global_noise.rds")
plot_model_pred_global_noise <- readRDS("../saves/plot_model_pred_global_noise.rds")

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  #geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = plot_model_pred_global_noise, aes(x=age+2,y = mean, col = model), size = 1)+
  geom_ribbon(data = plot_model_pred_global_noise, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = model), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(model + alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)+
  scale_fill_ptol(name = NULL)
```

@@

cor_prag <- model_pred%>%
  filter(model == "pragmatic",
         run == "no re")%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(run, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

cor_prag_re <- model_pred%>%
  filter(model == "pragmatic",
         run == "re")%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(run, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

cor_global <- model_pred%>%
  filter(model == "global")%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

cor_flat <- model_pred%>%
  filter(model == "flat")%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

cor_prior <- model_pred%>%
  filter(model == "prior")%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

cor_model_pred <- bind_rows(
  cor_prag,
  cor_prag_re,
  #cor_global,
  #cor_flat,
  #cor_prior
)

plot_cor_model_pred <- cor_model_pred %>%
  left_join(
    bind_rows(
      binnned_data%>%mutate(run = "no re"),
      binnned_data%>%mutate(run = "re"),
      # binnned_data%>%mutate(model = "flat"),
      # binnned_data%>%mutate(model = "prior")
      )
  )

@@

ggplot(data = plot_cor_model_pred,aes(x = model_mean, y = data_mean, col = alignment)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .5, alpha = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .5, alpha = .7)+
  geom_point(size = 1.5, stroke = 1, pch = 5)+
  coord_fixed()+
 stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mean, y = data_mean), inherit.aes = F, size = 3)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  facet_grid(run~age)+
  theme_few() + 
  scale_colour_ptol(name ="Alignment")
```

@@

>>>>

@@

x <- readRDS("../saves/sem_know_model_variant3_200k_model_comparison_noise.rds")

xx <- x %>%
  filter(!(Parameter %in% c("noise"))) %>%
  separate(Parameter, into = c("alignment","item", "age"), sep="/")%>%
  mutate(model = "prior_only_noise")%>%
  group_by(model,age,item, alignment) %>%
  summarise(mean = estimate_mode(value), 
            ci_lower = hdi_lower(value), 
            ci_upper =hdi_upper(value))%>%
  ungroup()%>%
  mutate(age = as.numeric(age))%>%
  left_join(aoa_ratings)%>%
  ungroup()%>%
  mutate(item = fct_reorder(factor(item), mean_aoa),
         source = "Model")

@@

### Pragmatic model

@@

model_pred_prag <- readRDS("../saves/model_pred_prag.rds")

plot_model_pred_prag <- model_pred_prag%>%
  group_by(alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = plot_model_pred_prag, aes(x=age+2,y = mean, col = alignment), size = 1)+
  geom_ribbon(data = plot_model_pred_prag, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol()+
  scale_fill_ptol()

x <- sample(1:length(unique(model_pred_prag$iteration)), 100)

plot_model_prag_line <- model_pred_prag%>%
  filter(iteration %in% x)

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_line(data = plot_model_prag_line, aes(x=age+2,y = pred, col = alignment, group = interaction(chain,iteration)),size = .2, alpha = .05)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "glm", method.args = list(family = "binomial"), se = T, alpha = .5, span = 2)+
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol(name = NULL)

@@

### Global pragmatic model

@@

model_pred_global<- readRDS("../saves/model_pred_global.rds")

plot_model_pred_global <- model_pred_global%>%
  group_by(alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

ggplot() +
  geom_hline(yintercept = 0.5, lty=2)+
  geom_smooth(data = data_comb_plot, aes(x = age_num, y = correct), col = "black", method = "loess", se = T, alpha = .5, span = 2)+
  geom_line(data = plot_model_pred_global, aes(x=age+2,y = mean, col = alignment), size = 1)+
  geom_ribbon(data = plot_model_pred_global, aes(x = age+2, ymin = ci_lower, ymax = ci_upper, fill = alignment), alpha = .4) +
  labs(x="Age",y="Mutual Exclusivity effect")+
  facet_grid(alignment~item)+
  theme_few() +
  ylim(-0.05,1.05)+
  xlim(2,5)+
  guides(alpha = F, fill = F)+ 
  scale_colour_ptol()+
  scale_fill_ptol()
```

@@

cor_prag <- model_pred_prag%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model,alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

# cor_prag_chain <- model_pred_prag%>%
#   mutate(age = as.numeric(age) + min(data_comb$age_num),
#          age = cut(age, 
#                       breaks = c(2,3,4,5),
#                       labels = c(2,3,4)))%>%
#   group_by(chain, alignment, item, age)%>%
#   summarise(model_mean = estimate_mode(pred),
#             model_ci_lower = hdi_lower(pred),
#             model_ci_upper = hdi_upper(pred))

cor_global <- model_pred_global%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

model_pred_flat<- readRDS("../saves/model_pred_flat.rds")

cor_flat <- model_pred_flat%>%
  mutate(age = as.numeric(age) + min(data_comb$age_num),
         age = cut(age, 
                      breaks = c(2,3,4,5),
                      labels = c(2,3,4)))%>%
  group_by(model, alignment, item, age)%>%
  summarise(model_mean = estimate_mode(pred),
            model_ci_lower = hdi_lower(pred),
            model_ci_upper = hdi_upper(pred))

model_pred_prior<- readRDS("../saves/model_pred_prior.rds")

# cor_prior <- model_pred_prior%>%
#   mutate(age = as.numeric(age) + min(data_comb$age_num),
#          age = cut(age, 
#                       breaks = c(2,3,4,5),
#                       labels = c(2,3,4)))%>%
#   group_by(model, alignment, item, age)%>%
#   summarise(model_mean = estimate_mode(pred),
#             model_ci_lower = hdi_lower(pred),
#             model_ci_upper = hdi_upper(pred))

cor_model_pred <- bind_rows(
  #cor_prag_chain,
  cor_prag,
  cor_global,
  cor_flat,
  cor_prior
)

plot_cor_model_pred <- cor_model_pred %>%
  left_join(
    bind_rows(
      # binnned_data%>%mutate(model = "pragmatic", chain = 1),
      # binnned_data%>%mutate(model = "pragmatic", chain = 2),
      # binnned_data%>%mutate(model = "pragmatic", chain = 3),
      # binnned_data%>%mutate(model = "pragmatic", chain = 4),
      # binnned_data%>%mutate(model = "pragmatic", chain = 5),
      # binnned_data%>%mutate(model = "pragmatic", chain = 6)
      binnned_data%>%mutate(model = "pragmatic"),
      binnned_data%>%mutate(model = "global"),
      binnned_data%>%mutate(model = "flat"),
      binnned_data%>%mutate(model = "prior")
      )
  )

@@

ggplot(data = plot_cor_model_pred,aes(x = model_mean, y = data_mean, col = alignment)) +
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = 1, size = .5)+
  geom_errorbar(aes(ymin = data_ci_lower, ymax = data_ci_upper),width = 0,size = .5, alpha = .7)+
  geom_errorbarh(aes(xmin = model_ci_lower, xmax = model_ci_upper), height = 0,size = .5, alpha = .7)+
  geom_point(size = 1.5, stroke = 1, pch = 5)+
  coord_fixed()+
 stat_cor(method = "pearson", label.x = 0.01, label.y = 0.99, aes(x = model_mean, y = data_mean), inherit.aes = F, size = 3)+
  xlim(0,1)+ylim(0,1)+
  xlab("Model")+
  ylab("Data")+
  facet_grid(model~age)+
  theme_few() + 
  scale_colour_ptol(name ="Alignment")+
  guides(col = F)
```

@@

plot_model_pred_flat <- model_pred_flat%>%
  group_by(alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

log_like_prag <- data_comb %>%
  select(-age)%>%
  mutate(age = age_month)%>%
  left_join(plot_model_pred_prag)%>%
  mutate(like = ifelse(correct == 1, mean, 1 - mean),
         log_like_prag = log(like))%>%
  select(age, alignment,item,log_like_prag)

log_like_global <- data_comb %>%
  select(-age)%>%
  mutate(age = age_month)%>%
  left_join(plot_model_pred_global)%>%
  mutate(like = ifelse(correct == 1, mean, 1 - mean),
         log_like_global = log(like))%>%
  select(age, alignment,item,log_like_global)

log_like_flat <- data_comb %>%
  select(-age)%>%
  mutate(age = age_month)%>%
  left_join(plot_model_pred_flat)%>%
  mutate(like = ifelse(correct == 1, mean, 1 - mean),
         log_like_flat = log(like))%>%
  select(age, alignment,item,log_like_flat)

model_comp <- log_like_prag%>%
  left_join(log_like_global)%>%
  left_join(log_like_flat)%>%
  mutate(prag_vs_global = log_like_prag - log_like_global,
         prag_vs_flat = log_like_prag - log_like_flat)%>%
  na.omit(prag_vs_global,prag_vs_flat)%>%
  summarize(prag_vs_global = logSumExp(prag_vs_global),
            prag_vs_flat = logSumExp(prag_vs_flat))

log_like <- bind_rows(
  readRDS("../saves/log_like_chain1.rds")%>%mutate(chain = 1),
  readRDS("../saves/log_like_chain2.rds")%>%mutate(chain = 2),
  readRDS("../saves/log_like_chain3.rds")%>%mutate(chain = 3),
  readRDS("../saves/log_like_chain4.rds")%>%mutate(chain = 4),
  readRDS("../saves/log_like_chain5.rds")%>%mutate(chain = 5)
)
